Data Predictive Control with Random Forests:\\Theory and Applications

### 2018-01-13 paper/final/intro.tex
small edits

\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
\textcolor[rgb]{0,0,1}{
\subsection{Related work}
This section is divided in 3 parts:
	\item In Section \ref{SSS:ModelBasedWork}
	\item In Section \ref{SSS:DataDrivenWork}
	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work with respect to the ones of our previous papers.
\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}
In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.
\subsubsection{Previous work}\label{SSS:PreviousWork}
In the following, the differences with respect to our previous papers are illustrated.
	\item In \cite{Behl2016} we	proposed an MPC-like control problem considering regression trees and ensemble methods algorithms for the Demand-Response problem. However, the proposed models allowed to optimally control the system considering only one-step lookahead prediction. Hence, it was not possible to use such approach to control the system make a prediction over an horizon of arbitrary length.
	\item This problem was addressed in \cite{Jain2016}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of a predictive horizon.
	This allowed us to setup an MPC-like control problem considering an arbitrary horizon.
	In this paper only single regression trees were considered, while ensemble methods were not took into account.
	However, modeling accuracy using single trees is strongly subject to the problems of the overfitting and high variance.
	For this reason, such approach has the advantage to be extremely simple from the complexity point of view, but the application is limited to specific kind of systems.
	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}, where an alternative approach is proposed to improve the system identification accuracy and robustness of the previous approach. More precisely, instead of considering a single tree with multiple output, we considered multiple trees with single output. Each tree is used to provide the prediction of the system's behavior over different steps of the horizon. With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results with respect to our previous work. In this paper we further extend this approach to one of the ensemble methods, i.e. random forests.
}

>>>>

\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.
\textcolor[rgb]{0,0,1}{
\subsection{Related work}
This section is divided in 3 parts:
\begin{enumerate}
	\item In Section \ref{SSS:ModelBasedWork} we provide a literature review of the approaches that make use of model-based techniques for energy optimization.
	\item In Section \ref{SSS:DataDrivenWork} we report the main works that use of data-driven methodologies for energy optimization and point out the difference with respect to our contribution.
	\item In Section \ref{SSS:PreviousWork} we underline the differences between the results presented in this work and the ones of our previous papers.
\end{enumerate}
\subsubsection{Model-based related work}\label{SSS:ModelBasedWork}
\subsubsection{Data-driven related work}\label{SSS:DataDrivenWork}}
In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

\textcolor[rgb]{0,0,1}{
\subsubsection{Previous work}\label{SSS:PreviousWork}
In the following, the differences with respect to our previous papers are illustrated.
\begin{itemize}
	\item In \cite{Behl2016} we	proposed an MPC-like control problem considering regression trees and ensemble methods algorithms for the Demand-Response problem. However, the proposed models allowed to optimally control the system considering only one-step lookahead prediction.
	Hence, it was not possible to use such approach to control the system considering a prediction over an horizon of arbitrary length.
	\item This problem was addressed in \cite{Jain2016}, where the regression trees algorithm was modified to allow the tree to provide multiple-output.
	Each output of the tree was then setup to provide the prediction of the system's behavior over different steps of the predictive horizon.
	This allowed us to setup an MPC-like control problem considering an arbitrary horizon.
	However, in this paper only single regression trees were considered, while ensemble methods were not took into account.
	Modeling accuracy using single trees is strongly subject to the problems of data overfitting and high variance.
	Such approach has the advantage to be extremely simple from the complexity point of view, but its application range is limited to specific kind of systems.
	\item In this work, we extend preliminary results gave in the conference paper \cite{JainCDC2017}.
	An alternative approach is proposed to improve the system's identification accuracy and robustness of the previous approach.
	More precisely, instead of considering a single tree with multiple output, we consider multiple trees with single output.
	Each tree is used to provide the prediction of the system's behavior over different steps of the horizon. With a small increase in complexity, we setup an MPC-like control problem with arbitrary horizon, that provides better results with respect to our previous work. In this paper we further extend this approach to one of the ensemble methods: the random forests.
\end{itemize}
}
